Supervised Reinforcement Learning with Partial States: \\ a Teaching Method for Action Selection for HRI

### 2017-07-24 paper.tex
Minor fixes

\maketitle
\begin{abstract}
%As robots will integrate human society, they will have to satify humans'
%expectations. However at the design time, it seems hardly possible to
%know in advance exactly all the behaviours expected from the robot. As such to
%be accepted in human society robots have to learn how to interact with humans.
    Learning social interactions is a complex task that robots will have to
    master before being accepted in society. However, as no simulator of human
    interactions sufficiently
    precise exists today, the learning have to be made online and in the real
    world. Actions executed by the robot have impact on humans, and as such
    have to be carefully selected, making the reliance on random exploration not
    applicable. Additionally, no clear reward function exists for
    social interactions. This implies that classical approaches such as
    Reinforcement Learning cannot be directly applied. As such we argue that
    robots will highly profit from human expertise and guidance to learn social
    interactions. And as the quantity of input a human can provide is limited,
    new methods have to be designed to use more efficiently human inputs. We
    combine the Supervised Progressively Autonomous Robot Competencies (SPARC)
    allowing a safer online learning with Reinforcement Learning based on
    partial states rather than full states to accelerate the learning by
    collecting and using more information from the human teacher.
\end{abstract}
%===============================================================================

@@

For a robot, interacting in the real world requires to combine inputs
 from many sensors to create a space composed of large number dimensions and
select actions appropriated to the current space. However,
 parts of the state can be irrelevant to the current goal and are not required
 to be taken into account to select the current action, only a limited number of
 features in the space are important. To interact efficiently, a robot has to
 learn to identify these features and associate them to related actions.

@@

The main framework for an agent to learn how to interact in an environment while
interacting is Reinforcement 
Learning (RL)~\cite{kober2013reinforcement,sutton1998reinforcement}. In RL, an
agent is interacting in an environment
providing numerical reward in reaction to the agent's actions and
the agent learns an action policy to maximise the expected cummulative discounted
reward obtained by the agent.

Classic approaches in RL generally rely on exploring virtual environments where the agent
tries to maximise the reward. In many cases where RL achieves success, the agent
has access to a virtual environment where the only real cost of exploring is 
time and energy, and the agent can interact as long as needed to gather
enough information on the environment to find a correct action policy. When
these conditions are reunited, RL can achieve impressive results, as shown by
the victory of a computer in Backgammon back in the
90s~\cite{tesauro1995temporal} or more recently the success mastering the game
of go~\cite{silver2016mastering}.

@@

\subsubsection{Design decisions:}
Initial knowledge has to be use in the designing of the algorithm, the
representation of the state and actions spaces and the reward funtion. For
example, only carefully crafted features allowed
Tesauro to improve its algorithm for Backgammon from a intermediate-level player
to super-human level~\cite{tesauro1995temporal}. Similarly, the design of
complex neural networks for state generalisation, the representation of actions
as motor primitive rather than raw motor angles or adding additional
informations in the reward function have important impacts on the ability of
the robot to reach a successful policy~\cite{kober2013reinforcement}.

\subsubsection{Demonstrations:}
Initial knowledge can be provided to the agent learning through demonstrations.
These demonstrations can be used to create an initial policy sufficiently
efficient to start interacting in the environment and gather information to
improve over time. This initial policy, required to receive meaningful feedback
from the environment can be impossible to reach by relying only on random
exploration and feedback from the environment. For example the game of Go in its larger board
size contains more than $10^{210}$ states, so exhaustive search is not possible.
Authors of \cite{silver2016mastering} started with
supervised learning from go master's games to learn a decent policy and then
used deep learning, self play and tree search to achieve super-human
capabilities and beating humans best players.

These demonstrations can also be used to learn a reward function. With Inverse
Reinforcement Learning, the agent is not provided with a reward function, but
derives it from a set of expert demonstrations. The agent can then explore
around the demonstrated policy to find how to optimise the reward function. With
this approach, Abbeel and Ng achieved better than human control for a robotic
helicopter~\cite{abbeel2004apprenticeship} based on demonstrations from experts.

@@

A first approach consists in sequencing tasks the robot will face. 
This approach, known as "scaffolding" \cite{saunders2006teaching}, progressively
increase the difficulty and complexity of the task as the robot is learning
reaching policies not directly learnable in the final environment.

A second approach is "shaping". For example, the TAMER framework
\cite{knox2009interactively}, allows to teach an agent an action policy in an
environment providing no rewards using a human observer to supply rewards to the
agent. With this approach the teacher can adapt the
rewards to the current policy and performance of the agent, for example starting to reward
positively suboptimal behaviours but better than the current policy and
later, as the agent improve its behaviour, reward these same suboptimal
behaviours negatively and only the optimal ones positively.

@@

In \cite{senft2015sparc}, we introduced the Supervised Progressively
Autonomous Robot Competencies (SPARC). SPARC relies on supervisor with the
ability to control the robot's actions. This teacher is presented with the
actions the robot is about to execute and can cancel an action, let it be
executed or select another to be executed. Based on the supervisor's decisions,
the robot learns which actions are desired and can refine its policy over time
reducing the need for the teacher to correct or select an action.

SPARC gives control on the robot actions to an expert, who can guide the
exploration in the desired part of the environment and this ensures the robot's
behaviour is consistently appropriate. As the exploration is
guided, and all the actions are useful, the learning can be faster than
autonomous learning or feedback based learning as illustrated in Figure
\ref{fig:comparison}.

@@

In \cite{senft2017supervised}, we presented a way to combine SPARC and RL, by
assigning a positive reward to every action executed by the robot as it would
have been explicitely or implicitely approved by the supervisor. However this method
was not making use of any kind of generalisation, directly mapping a
single (state, action) pair to a reward, and as such is not applicable to
environments with a continuous or high dimensional space or undeterministic
transition from one state to another.

@@

To make RL applicable in high dimensional or continuous states, the algorithm
requires a way to generalise knowledge to unseen states. 
A classical approach is to use a first layer of neural
network or to use deep reinforcement learning relying on deep neural
networks to learn a mapping state action Qvalue for example. 
These approaches, neural network based, relies on having a large number of
datapoints to converge toward a good function approximator. However, 
 in many application, especially HRI, these amounts of data are not
obtainable due to time or unconsistency and methods to generalise with a lower
number of inputs are required.

@@

In this paper we propose to use a human to highlight the relevant features of
the environment to reduce the state dimension of the points stored in memory.
We introduce the concept of partial state, a sliced version of the state defined
only on a subset of the dimensions of the state. This shifts the (state,action)
pair paradigm to (partial state,action)  
and in the case of RL, the tuple (state, action, reward) to (partial state,
action, reward) and as such allows to compare the current state and the
datapoints only on relevant features for action selection.

@@

As the supervisor can estimate the future impacts of an action, the problem of
credit assignment for delayed rewards can be ignored  and allows us to consider
only a myopic approach.

For this paper, we will reuse the formalism of rewardless Markov Decision 
Process to identify the different elements of our system. The agent has access
to a set of actions $\mathcal{A}$, a state $\mathcal{S} \in [0,1]^{n}$. We
also define the partial states $\mathcal{S'} \in [0,1]^{n'}$ with $n' \leq n$ as a
slice of $\mathcal{S}$, a subset of $\mathcal{S}$ where some dimensions have
been removed. 
%With each dimension of S representing features in the environment.

At the instant $t$, the agent has access the state $s$, an action $a$, evaluated
with a reward $r$ associated to the partial state $s'$. For example, a state $s$ could be
defined in 4 dimensions such as ${s=[1,0.2,0,0.5]}$, and $s'$ in two dimensions with
${s'=[-,0.2,0,-]}$ with symbol '$-$' reprensenting the dimensions removed. The
action $a$ has been evaluated $r$ in the partial state $s'$.

To each action
$a \in \mathcal{A}$ we can associate a set $\mathcal{C}_{a}$ of pairs $(s',r)$
representing the rating done by the supervisor to action $a$ with feature
highlighted for $s'$. When adding a new pairt $(s',r)$, we can discard potential
previous pairs with an identical $s'$ to represent the evolution of the policy
evaluation by the supervisor.

@@

    \caption{Algorithm for selecting an action based on previous
    partial state action rewards tuples and current state}
    \label{algo}
\end{algorithm}

@@

SPARC has been showed to be compatible with RL in \cite{senft2017supervised},
and can also be easily combined with the approach presented in this paper using
partial states. For example, when selecting an action for the robot to be
executed, the supervisor can also select which features in the environment
should be selected in the partial state, and this associates the reward +1
to this action in this partial state. Similarly when an action is
proposed to the supervisor the features represented by the dimensions of the
closest partial state can be exposed to the supervisor as a way to explain why
this action has been selected. Facing this, the supervisor can not react,
letting the action being executed and associating a reward of +1 to the partial
state and the action proposed, change the partial state to correct the features
related to this action or cancel it, preventing the execution and associating a
reward of -1.

@@

An example of application yet to be tested is robot for teaching to children.
A robot can play an educative game with children to teach them notions about
diverse topics according to the needs of the teacher. For this example, a child
and a robot are playing a game about food web, teaching wich animals eat which
ones on a Sandtray~\cite{baxter2012touchscreen} as shown in
Figure~\ref{fig:game}. In addition to the child and
the robot, an adult supervises the robot using a tablet with a Graphical User
Interface (GUI) (Figure~\ref{fig:gui}) to teach the robot how to interact with the child.
\begin{figure}
        \centering
  \includegraphics[width=80mm]{./fig/game.png} 
    \caption{Game displayed on the Sandtray the child is playing with the robot.}
        \label{fig:game}
\end{figure}

@@

The main limitations of the approach reside in the difference of
representation of the state and action spaces between the supervisor and the
algorithm and the limit in communication. For example a user could try to move
an animal close to another one,
and depending of the represenation of the actions on the algorithm side, the
action might not be understood in the same way. Similarly, features used by the
supervisor to select actions might not be represented in the state used by the
algorithm. And lastly, in the case of implicit selection of features, a single
case of features representatiion (for example highlighting two animals) might not
be perceived in the same way by observers.

@@

The current work could also be extended to allow the agent continue to improve its
behaviour even in the absence of supervisor, progressively exploring around the
learnt policy improving its behaviour beyond the demonstrations. This
could be done by allowing the supervisor to provide rewards during the learning
phase, and combine these rewards with the demonstrations to learn a reward
function in a fashion similar to Inverse Reinforcement Learning
\cite{abbeel2004apprenticeship}. These could also use partial states associated
to these rewards to ease the generalisation of the reward function with only a
low number of datapoints.
But without the supervisor, the assumption that only a
myopic action selection is sufficient would not hold anymore and the problem of
delayed rewards would have to be tackled.
%===============================================================================
%\section{Conclusion}
\label{sec:conclusion}
%==============================================================================
\section{Acknowledgments}
This work was supported by the EU FP7 DREAM project (grant no.  611391) and EU
H2020 Marie Sklodowska-Curie Actions project DoRoThy (grant 657227).

>>>>

\maketitle
\begin{abstract}
%As robots will integrate human society, they will have to satify humans'
%expectations. However at the design time, it seems hardly possible to
%know in advance exactly all the behaviours expected from the robot. As such to
%be accepted in human society robots have to learn how to interact with humans.
    Learning social interactions is a complex task that robots will have to
    master before being accepted in society. However, as no simulator of human
    interactions sufficiently
    precise exists today, the learning have to be made online and in the real
    world. Actions executed by the robot have impacts on humans, and as such
    have to be carefully selected, making the reliance on random exploration not
    applicable. Additionally, no clear reward function exists for
    social interactions. This implies that classical approaches such as
    Reinforcement Learning cannot be directly applied for learning how to
    interact with humans. As such we argue that
    robots will highly profit from human expertise and guidance to learn social
    interactions. And as the quantity of input a human can provide is limited,
    new methods have to be designed to use more efficiently human inputs. We
    combine the Supervised Progressively Autonomous Robot Competencies (SPARC)
    allowing a safer online learning with Reinforcement Learning and use
    partial states rather than full states to accelerate generalisation and
    obtain more quickly a usable action policy.
\end{abstract}
%===============================================================================

@@

For a robot, interacting in the real world requires to combine inputs
 from many sensors to create a space composed of large number dimensions and
select actions appropriated to the current space. However,
 parts of the state can be irrelevant to the current goal or current state and
 are not required
 to be taken into account to select the current action, only a limited number of
 features in the space are important. To interact efficiently, a robot has to
 learn to identify these features and associate them to related actions.

@@

The main framework for an agent to learn how to interact in an environment while
interacting is Reinforcement 
Learning (RL)~\cite{kober2013reinforcement,sutton1998reinforcement}. In RL, an
agent is interacting in an environment
providing numerical reward in reaction to the agent's actions and
the agent learns an action policy to maximise the expected cummulative discounted
reward received.

%Classic approaches in RL generally rely on exploring virtual environments where the agent
%tries to maximise the reward.
In many cases where RL achieves success, the agent
has access to a virtual environment where the only real cost of exploring is 
time and energy, and the agent can interact as long as needed to gather
enough information on the environment to find a correct action policy. When
these conditions are reunited, RL can achieve impressive results, as shown by
the victory of a computer in Backgammon back in the
90s~\cite{tesauro1995temporal} or more recently the success mastering the game
of go~\cite{silver2016mastering}. Even where a virtual environment is available,
but especially when it is not, human knowledge is required at many steps of the
design of the algorithm or the learning phase to achieve an efficient action
policy.

@@

\subsubsection{Design decisions:}
Initial knowledge has to be use in the designing of the algorithm, the
representation of the state and actions spaces and the reward funtion. For
example, only carefully crafted features allowed
Tesauro to improve its algorithm for Backgammon from a intermediate-level player
to super-human level~\cite{tesauro1995temporal}. Similarly, the design of
complex neural networks for state generalisation, the representation of actions
as motor primitives rather than raw motor angles or adding additional
information in the reward function have important impacts on the ability of
the robot to reach a successful policy~\cite{kober2013reinforcement}.

\subsubsection{Demonstrations:}
Initial knowledge can be provided to the agent learning through demonstrations.
These demonstrations can be used to create an initial policy sufficiently
efficient to start interacting in the environment and gather information to
improve over time. This initial policy, required to receive meaningful feedback
from the environment can be impossible to reach by relying only on random
exploration and feedback from the environment. For example the game of Go in its larger board
size contains more than $10^{210}$ states, so exhaustive search is not possible.
Authors of \cite{silver2016mastering} started with
supervised learning from go masters' games to learn a decent policy and then
used deep learning, self play and tree search to achieve super-human
capabilities and beating human best players.

These demonstrations can also be used to learn a reward function. With Inverse
Reinforcement Learning, the agent is not provided with a reward function, but
derives it from a set of expert demonstrations. The agent can then explore
around the demonstrated policy to optimise the reward function. With
this approach, Abbeel and Ng achieved better than human control for a robotic
helicopter~\cite{abbeel2004apprenticeship} based on demonstrations from experts
and further exploration and autonomous learning.

@@

A first approach consists in sequencing the tasks the robot will face. 
This approach, known as "scaffolding" \cite{saunders2006teaching}, progressively
increases the difficulty and complexity of the task as the robot is learning
reaching policies not directly learnable in the final environment.

Approaches from animal conditioning have also been transfered to Machine Learning. 
For example, the TAMER framework~\cite{knox2009interactively} is an adaptation of "shaping",
tuning an animal's behaviour by providing rewards to guide it to the desired
behaviour. Unlike other RL setups, with TAMER, the
reward is not provided by the environment but by a human observer. With this
approach the teacher can adapt the 
rewards to the current policy and performance of the agent, for example starting to reward
positively suboptimal behaviours but better than the current policy and
later, as the agent improve its behaviour, reward these same suboptimal
behaviours negatively and only the optimal ones positively.

@@

In \cite{senft2015sparc}, we introduced the Supervised Progressively
Autonomous Robot Competencies (SPARC). SPARC relies on a supervisor with the
ability to control the robot's actions. This teacher is presented with the
actions the robot is about to execute and can cancel an action about to be
executed, let it be
executed or select another to be executed. Based on the supervisor's decisions,
the robot learns which actions are desired and can refine its policy over time
reducing the need for the teacher to correct or select an action.

SPARC gives control on the robot actions to an expert, who can guide the
exploration in the desired part of the environment which ensures the robot's
behaviour is consistently appropriate. As the exploration is
guided, and all the actions are useful, the learning can be faster than
autonomous learning or feedback based learning as illustrated in Figure
\ref{fig:comparison}.

@@

In \cite{senft2017supervised}, we presented a way to combine SPARC and RL, by
assigning a positive reward to every action executed by the robot as it would
have been explicitely or implicitely approved by the supervisor. However this method
was  directly mapping a single (state, action) pair to a reward without making
use of any kind of generalisation, it was not applicable to
environments with a continuous or high dimensional space or undeterministic
transition from one state to another, such as interactions with humans.

@@

To make RL applicable in high dimensional or continuous states, the algorithm
requires a way to generalise knowledge to unseen states. 
A classical approach is to use a first layer of neural
network or to use deep reinforcement learning relying on deep neural
networks to learn a mapping state action Qvalue for example. 
These approaches, neural network based, relies on having a large number of
datapoints to converge toward a good function approximator. However, 
 in many application, especially in HRI, these amounts of data are not
obtainable due to time or unconsistency of human reactions, so methods to
generalise with a lower number of inputs are required.

@@

In this paper we propose to use a human to highlight the relevant features of
the environment to reduce the state dimension of the points only to relevant
information.
We introduce the concept of partial state, a sliced version of the state defined
only on a subset of the dimensions of the state. This shifts the (state, action)
pair paradigm to (partial state, action)  
and in the case of RL, the tuple (state, action, reward) to (partial state,
action, reward). This allows to compare the current state and the
datapoints only on relevant features for action selection.

@@

As the supervisor can estimate the future impacts of an action, the problem of
credit assignment for delayed rewards can be ignored which allows us to consider
only a myopic approach.

For this paper, we will reuse the formalism of rewardless Markov Decision 
Process to identify the different elements of our system. The agent has access
to a set of actions $\mathcal{A}$ and a state $\mathcal{S} \in [0,1]^{n}$. We
also define the partial states $\mathcal{S'} \in [0,1]^{n'}$ with $n' \leq n$ as a
slice of $\mathcal{S}$, a subset of $\mathcal{S}$ where some dimensions have
been removed. 
%With each dimension of S representing features in the environment.

When the agent execute an action $a$ in state $s$, it receives the reward $r$
associated with the partial state $s'$. For example, a state $s$ could be
defined in 4 dimensions such as ${s=[1,0.2,0,0.5]}$, and $s'$ in two dimensions with
${s'=[-,0.2,0,-]}$ with symbol '$-$' reprensenting the dimensions removed. For
the learning algorithm, this means that the action $a$ has been evaluated $r$ in
the partial state $s'$.

To each action
$a \in \mathcal{A}$ we can associate a set $\mathcal{C}_{a}$ of pairs $(s',r)$
representing the rating done by the supervisor to action $a$ with features
highlighted for the multiple $s'$. When adding a new pairt $(s',r)$, we can discard potential
previous pairs with an identical $s'$ to represent the evolution of the policy
evaluation by the supervisor.

@@

    \caption{Algorithm for selecting an action based on previous
    partial state action rewards tuples and current state.}
    \label{algo}
\end{algorithm}

@@

SPARC has been showed to be compatible with RL in \cite{senft2017supervised},
and can also be easily combined with the approach presented in this paper using
partial states. For example, when selecting an action for the robot to be
executed, the supervisor can also select which features in the environment
should be selected in the partial state, and this associates the reward +1
to this action in this partial state. Similarly when an action is
proposed to the supervisor the features represented by the dimensions of the
closest partial state can be exposed to the supervisor as a way to explain why
this action has been selected. Facing this, the supervisor can: not react,
letting the action being executed and associating a reward of +1 to the partial
state and the action proposed, change the partial state to correct the features
related to this action or cancel it, preventing the execution and associating a
reward of -1 to the partial state identified by the robot or the supervisor.

@@

An example of application yet to be tested is robot for teaching children.
A robot can play an educative game with children to teach them notions about
diverse topics according to the needs of the teacher. For this example, a child
and a robot are playing a game about food web, teaching wich animals eat which
ones on a Sandtray~\cite{baxter2012touchscreen} as shown in
Figure~\ref{fig:game}. In addition to the child and
the robot, an adult supervises the robot using a tablet with a Graphical User
Interface (GUI) (Figure~\ref{fig:gui}) to teach the robot how to interact with the child.
\begin{figure}
        \centering
  \includegraphics[width=80mm]{./fig/game.png} 
    \caption{Game displayed on the Sandtray the child is playing with the robot.}
        \label{fig:game}
\end{figure}

@@

The main limitations of the approach reside in the difference of
representation of the state and action spaces between the supervisor and the
algorithm and the limit in communication. For example a user could try to move
an animal close to another one,
and depending of the representation of the actions on the algorithm side, the
action might not be understood in the same way. Similarly, features used by the
supervisor to select actions might not be represented in the state used by the
algorithm. And lastly, in the case of implicit selection of features, a single
case of features representation (for example highlighting two animals) might not
be perceived in the same way by observers.

@@

The current work could also be extended to allow the agent continue to improve its
behaviour even in the absence of a supervisor, progressively exploring around the
learnt policy improving its behaviour beyond the demonstrations. This
could be done by allowing the supervisor to provide rewards during the learning
phase and combine these rewards with the demonstrations to learn a reward
function in a fashion similar to Inverse Reinforcement Learning
\cite{abbeel2004apprenticeship} or TAMER \cite{knox2009interactively}. This
could also use partial states associated to these rewards to ease the
generalisation of the reward function with only a low number of datapoints.
But without the supervisor, the assumption that only a
myopic action selection is sufficient would not hold anymore and the problem of
delayed rewards would have to be tackled.
%===============================================================================
%\section{Conclusion}
%\label{sec:conclusion}
%==============================================================================
\section{Acknowledgments}
This work was supported by the EU FP7 DREAM project (grant no.  611391) and EU
H2020 Marie Sklodowska-Curie Actions project DoRoThy (grant 657227).
